---
title: "WWII Eastern Theater Preliminary Map Ideas"
date: "September 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Synopsis
I took Romerl's data loading and cleaning and kept the lattitude and longitude columns.  Using ggplot2, maps, and mapdata, I plotted the location of bombing sites across Germany.  The graphics are very preliminary, this is to suggest a form of data visualization we could use, not to present anything that would go in the final presentation. 

## Summary
```{r load-libraries, eval=TRUE}
library(stringr)
library(ggplot2)
library(maps)
library(mapdata)
library(ggrepel) #not using this at the moment, but it does give the option to add labels.  While not useful for the bombing sites, we could use it to label cities.
```

Loading of THOR_WWII_DATA_CLEAN.csv directly from data.world website.

```{r load-data, eval=TRUE}
# ref: [HOW]
urlfile <- "https://query.data.world/s/7tdvewopqdr5mu4zwlqeuy4c7nyeco"
tableWW2 <- read.table(file = urlfile, header = TRUE, fill = TRUE, sep = ",")
```



Removing unnecessary fields.

```{r cleandataframe, eval=TRUE, include=FALSE}
tableWW2 <- tableWW2[, -which(names(tableWW2) %in% c("TGT_COUNTRY_CODE","TGT_ID","TGT_INDUSTRY_CODE","SOURCE_LATITUDE","SOURCE_LONGITUDE","MDS","TAKEOFF_LATITUDE","TAKEOFF_LONGITUDE","TARGET_COMMENT","MISSION_COMMENTS","SOURCE","DATABASE_EDIT_COMMENTS", "BDA", "CALLSIGN", "ROUNDS_AMMO", "SPARES_RETURN_AC","WX_FAIL_AC", "MECH_FAIL_AC", "MISC_FAIL_AC", "TIME_OVER_TARGET","SIGHTING_METHOD_CODE","SIGHTING_EXPLANATION"))]
```

Separating data frame into two theaters: European and Pacific

```{r separatingdataframe, eval=TRUE, include=FALSE}
tableWW2ETO <- subset(tableWW2, THEATER == "ETO")
tableWW2PTO <- subset(tableWW2, THEATER == "PTO")

```

##Map stuff

```{r}
ETOsites<-tableWW2ETO[,9:12]
ETOclean<-ETOsites[complete.cases(ETOsites),]
Targets<-ETOsites$TGT_TYPE
wm<-map_data("world")
sites<-data.frame(lat=ETOclean$LATITUDE,long=ETOclean$LONGITUDE, stringsAsFactors = FALSE)
#ggplot(ETOclean,aes(x=LONGITUDE,y=LATITUDE))+geom_point()+scale_y_continuous(limits=c(30,75))+scale_x_continuous(limits=c(-5,30))
EuropeM<-ggplot() + geom_polygon(data = wm, aes(x=long, y = lat, group = group), fill = NA, color = "black") + coord_fixed(1.3)
EuropeM<-EuropeM+xlim(-10,40)+ylim(35,60)
```


```{r}
#f50k<-sites[1:50000,]
EuropeM+geom_point(data=sites,aes(long,lat),color="red",size=1)
```

```{r}
justgerm<-ggplot() + geom_polygon(data = wm, aes(x=long, y = lat, group = group), fill = NA, color = "black") + coord_fixed(1.3)+xlim(2,16)+ylim(46,55)
justgerm+geom_point(data=sites,aes(long,lat),color="red",size=1)
```

```{r}

Airfield<-subset(ETOclean,str_detect(ETOclean$TGT_TYPE,"AIR"))
City<-subset(ETOclean,str_detect(ETOclean$TGT_TYPE,"CITY"))
oil<-subset(ETOclean,str_detect(ETOclean$TGT_TYPE,"OIL"))
Railyard<-subset(ETOclean,ETOclean$TGT_TYPE=="MARSHALLING YARD")
EuropeM+geom_point(data=City,aes(LONGITUDE,LATITUDE),color="black",size=1)+geom_point(data=Airfield,aes(LONGITUDE,LATITUDE),color="red",size=1)+geom_point(data=Railyard,aes(LONGITUDE,LATITUDE),color="blue",size=1)+geom_point(data=oil,aes(LONGITUDE,LATITUDE),color="green",size=1)
justgerm+geom_point(data=Airfield,aes(LONGITUDE,LATITUDE),color="red",size=1)+geom_point(data=Railyard,aes(LONGITUDE,LATITUDE),color="blue",size=1)+geom_point(data=oil,aes(LONGITUDE,LATITUDE),color="green",size=1)+geom_point(data=City,aes(LONGITUDE,LATITUDE),color="black",size=1)

```
